lifespan_eda

### 2024-01-22 eda/lifespan_eda.Rmd
fixing sites with the same name that combine across studies

vol_list_global <- readRDS(file=paste0(base_path, "/cubic_scripts/R_scripts/vol_list_global.rds"))

base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source(paste0(base_path, "/cubic_scripts/R_scripts/gamlss_helper_funs.R"))
```

@@

## Site-Level Data
look just at data w site info

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site))) %>%
  na.omit()
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

@@

```{r}
ggplot(cn.df.site) +
  geom_violin(aes(x=study, y=age_days), scale="width")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.site) +
  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

@@

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site))) %>%
  na.omit()

unique(cn.df.imp$site)
length(is.na(cn.df.imp$site))
max(cn.df.imp$age_days)/365.25
mean(cn.df.imp$age_days)/365.25
```

```{r}
ggplot(cn.df.imp) +
  geom_violin(aes(x=study, y=age_days), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.imp) +
  geom_violin(aes(x=site, y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

>>>>

base_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar"
vol_list_global <- readRDS(file=paste0(base_path, "/cubic_scripts/R_scripts/vol_list_global.rds"))

source("/Users/megardn/Desktop/BGD_Repos/GAMLSS/plotting_functions.R")
source(paste0(base_path, "/cubic_scripts/R_scripts/gamlss_helper_funs.R"))
```

@@

## Site-Level Data

look just at data w site info

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site))) %>%
  unite(site, study, site, sep = "_", remove = FALSE) %>% #so sites with the same name don't get combined across studies!
  na.omit()
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

@@

```{r}
ggplot(cn.df.site) +
  geom_boxplot(aes(x=reorder(study, age_days), y=age_days))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.site) +
  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```

@@

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="_")))) %>%
  na.omit()

unique(cn.df.imp$site)
table(cn.df.imp$site, cn.df.imp$study)
sum(is.na(cn.df.imp$site))
max(cn.df.imp$age_days)/365.25
mean(cn.df.imp$age_days)/365.25
```

```{r}
ggplot(cn.df.imp) +
  geom_violin(aes(x=reorder(study, age_days), y=age_days), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(cn.df.imp) +
  geom_violin(aes(x=reorder(site, age_days), y=age_days, fill=study), scale="width") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position = "none")
```
